Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2018-07-30 paper/Cortex_submission/darkcontrol_paper.tex
new Cortex

% \title{Dark Control:\\
%        Towards a Unified Account of Default Mode Function\\
%        by Markov Decision Processes}
\title{Dark Control: The Default Mode Network as a Reinforcement Learning Agent}

@@

\begin{abstract}
The default mode network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its highest energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an overarching function.
%
Many research streams speak in favor of an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
The main purpose of the DMN, we argue,
may be to perform optimization based on Markov Decision Processes
through vicarious trial and error.
%
Our formal account of DMN function naturally accommodates as special cases
previous interpretations based on
(1) predictive coding,
(2) semantic associations, and
(3) a sentinel role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

The present work adopts the perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes
of really happened, hypothetically imagined, and
expected futures to optimize behavioral performance.
Formally, we propose reinforcement learning to be a particularly attractive framework
to describe, quantify, DMN function and the brain.
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so partially determine their next stimuli, in particular stimuli that are necessary for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

@@

Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across multiple time-scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
optimize the behavior of the organism.

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing some of the first empirical evidence that
morphological variability in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments.
At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
Notably, our account does not claim that the brain or neural activity in the DMN are completely identical with reinforcement learning algorithms. Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to understand DMN function.

@@

\section{Reinforcement learning control: a process model for DMN function}
We argue the outlined neurobiological properties
of the DMN regions
to be sufficient for implementing all components
of a full-fledged reinforcement learning (RL) system.
Recalling past experience, considering candidate actions,
random sampling of possible experiences, as well as
estimation of instantaneous and expected delayed reward outcomes
are key components of intelligent RL agents
that are plausible to functionally intersect in the DMN.

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time-scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. An even recent work~\citep{robinowitz2018}
develops an MDP-based reasoning about future behavior
of other agents (the \textit{ToM-Net}).
The idea is to use meta-learning (learning to learn) to build strong priors
about the behavior of of a population of other agents.

% As such, we aim to expand MDP applications as a useful model from "online" decision-making to the realms of "offline" behaviors most associated with the DMN.

@@

\paragraph*{Reinforcement learning in the brain.}
RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
Indeed,
it has been proposed \citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a mapping between algorithmic aspects
underlying model-free and model-based RL learning
and neurobiological aspects underlying brain function.
The neurotransmitter dopamine could serve
as a 'teaching signal' to better estimate value associations
and action policies by modulating
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
reward contingencies, etc.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

% In contrast, model-free methods require no pre-specified knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof. The agent infers which state-action pairs lead to reward through sampling the world in a trial-and-error manner and derives longer-term reward aggregates using environmental feedback information as an incentive. In so doing, model-free agents ultimately learn both an action policy and an implicit representations of the world. This distinction between model-free and model-based RL is similar to previous views \citep{dayan2014}.

@@

The goal of the RL agent is then to successively update this action policy in order
to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
the definition of cumulative reward $G^\pi$,
the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
viewed to be characteristic for a certain agent.
On the one hand,
setting $\gamma = 0$ yields perfectly hedonistic behavior.
An agent with such a shortsighted time horizon is exclusively
concerned with immediate rewards.
This is however not compatible with coordinated planning of longer-term agendas that is
potentially subserved by neural activity in the DMN.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
On the other hand,
setting $0 < \gamma < 1$ allows a learning process to arise.
A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
that is, the behavioral predispositions related to trading longer delays
for higher reward outcomes.
Such an agent puts relatively more emphasis on rewards expected in
a more distant future.
Concretely,
rewards that are not expected to occur only within
% \footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
a very large number of
time steps from the present point are ignored.
The complexity reduction by time discounting
alleviates the variance of expected rewards accumulated across
considered action cascades by limiting the depth of the search tree.
Given that there is more uncertainty in the far future,
it is important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

@@

\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
In RL, Monte Carlo simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
Monte Carlo simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
actions and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
These simulations of experience batches
would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
A brain-imaging experiment in humans for instance identified hippocampal signals that specifically preceded upcoming choice performance in prospective planning in new environments
\citep{kaplan2017neural}.
It would be a simple strategy to evaluate
all legal actions and selecting the action with
highest expected cumulative rewards.
In MDPs, MC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
This is because expected consequences of action choices can be well evaluated although only a subset of the states are actually considered \citep{daw2014algorithmic}.
% Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?}

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to
incorporate the previous states of such model into the current state. This
extension may partially account for implicit memory at the behavioral level, but
may not explain the underlying neurobiological implementation or accommodate
explicit memory. Implicit memory-based processing arises in our MDP account of
DMN function in several different forms: successive updates of a) the action
policy and the value function, both being products of the past, as well as b)
the deep non-linear relationships within the hierarchical connections of
biological neural networks. The brain's adaptive synaptic connections can be
thought as a deep neural-network architecture affording an implicit form of
information compression of life experience. Such memory traces are stored in the
neural machinery and can be implicitly retrieved as a form of knowledge during
simulation of action rather than accessed as a stored explicit representation
\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
be seen as some type of Monte Carlo sampling for memory recall, which can also
be a basis for probabilistic simulations across time-scales
\citep{schacter2007remembering,Axelrod2017}.

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context. In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing intervention of the organism on the world in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.

@@

From the perspective of a RL agent,
prediction in the DMN reduces to
generalization of
policy and value computations from sampled experiences to
successful action choices and reward predictions in future states.
As such,
plasticity in the DMN arises naturally.
If an agent behaving optimally in a certain environment moves
to new, yet unexperienced environment, reward prediction errors will
largely increase.
This feedback will lead to adaptation of policy considerations
and value estimations until the intelligent system converges to a
new steady state of optimal action decisions in a volatile world.

@@

\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits
\citep{luksys2009stress, haushofer2014psychology}.
Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation. Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.

@@

In sum,
recasting DMN function in terms of MDPs therefore naturally incorporates
a majority of aspects from the prediction coding hypothesis.
The present MDP account of DMN function may therefore
serve as a concrete implementation of
predictive coding ideas.
MDPs have the advantage of exposing an explicit
mechanisms for modulating the horizon of future considerations and for
how the internal representation of the world is updated,
as well as why certain predictions may be more relevant to the agent than others.

@@

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be explored.
In both accounts,
there is also little reason to assume that contemplating alternative realities of various
levels of complexity, abstraction, time-scale, and purpose
rely on mechanisms that are qualitatively different. This interpretation concurs with
DMN activity increases across time, space, and content domains
demonstrated in many brain-imaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account offer explanations why HC damage does
not only impair recalling past events, but also imagining hypothetical and future
scenarios \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-enabled, internally
generated information for probabilistic representation of action outcomes,
MDPs render explicit the grounds on which an action is
eventually chosen, namely, the estimated cumulative reward.
In contrast to many versions of the semantic hypothesis,
the MDPs naturally integrate the egocentric view
(more related to current action, state, and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not provide
suffcient explanation of \textit{how}
explicit world knowledge and logical analogies thereof
lead to foresight of future actions and states.
The semantic hypothesis does also not fully explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space exploration
by Monte Carlo approximation achieves better fine-tuning of the
action policies and inference of expected reward outcomes.
That is, the purposeful stochasticity of policy and value updates
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, mental scene construction according to the semantic
account is lacking an explicit time and incentive model,
both of which are integral parts of the MDP interpretation of DMN function.

>>>>

\title{Dark Control:\\
       Towards a Unified Account of Default Mode Function\\
       by Markov Decision Processes}

@@

\begin{abstract}
The default mode network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its highest energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an overarching function.
%
Many research streams speak in favor of an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
Specifically, we answer the question whether the neurobiological properties of the DMN
collectively provide the computational building blocks necessary for a Markov Decision Process.
%
We argue that our formal account of DMN function naturally accommodates as special cases
previous interpretations based on
(1) predictive coding,
(2) semantic associations, and
(3) a sentinel role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

The present work adopts the perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes
of really happened, hypothetically imagined, and
expected futures to optimize behavioral performance.
Formally, a particularly attractive framework
to describe, quantify, and predict intelligent systems, such as the brain,
is proposed to be the combination of control theory and
reinforcement learning (RL). 
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so partially determine their next stimuli, in particular stimuli that are necessary for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

@@

Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across multiple time-scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
Even (unavoidably) imperfect memory recall,
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
optimize the behavior of the organism.

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing some of the first empirical evidence that
morphological variability in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments.
At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.

@@

\section{Reinforcement learning control: a process model for DMN function}
We argue the outlined neurobiological properties
of the DMN regions
to be sufficient for implementing all components
of a full-fledged reinforcement-learning (RL) system.
Recalling past experience, considering candidate actions,
random sampling of possible experiences, as well as
estimation of instantaneous and expected delayed reward outcomes
are key components of intelligent RL agents
that are plausible to functionally intersect in the DMN.

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time-scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation of MCMC sampling.

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models are applicable to something as complex as human behavior. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has successfully used MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics \citep{ng2004,abbeel2004} and in challenging multistep strategy games ~\citep{mnih2015,silver2016mastering,pritzel2017neural}.
%
As such, we aim to expand MDP applications as a useful model from "online" decision-making to the realms of "offline" behaviors most associated with the DMN.

@@

\paragraph*{Towards model-free reinforcement learning for the DMN.}
Model-free RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
Indeed,
it has been proposed \citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a direct mapping between model-free RL learning algorithms
and aspects of the brain.
The neurotransmitter dopamine could serve
as a 'teaching signal' to better estimate value associations
and action policies by controlling
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
reward contingencies, etc.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

In contrast, model-free methods require no prespecified knowledge of the environment (transition probabilities, types of sensory input, etc.) or representation thereof. The agent infers which state-action pairs lead to reward through sampling the world in a trial-and-error manner and derives longer-term reward aggregates using environmental feedback information as an incentive. In so doing, model-free agents ultimately learn both an action policy and an implicit representations of the world. This distinction between model-free and model-based RL is similar to previous views \citep{dayan2014}.

@@

The goal of the RL agent is then to successively update this action policy in order
to maximize $G^\pi$ on average (cf. below). In \eqref{eq:cumr},
the definition of cumulative reward $G^\pi$,
the constant $\gamma$ $(0 \le \gamma < 1$) is the \textit{reward discount factor},
viewed to be characteristic for a certain agent.
On the one hand,
setting $\gamma = 0$ yields perfectly hedonistic behavior.
An agent with such a shortsighted time horizon is exclusively
concerned with immediate rewards.
This is however not compatible with coordinated planning of longer-term agendas that is
potentially subserved by neural activity in the DMN.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
On the other hand,
setting $0 < \gamma < 1$ allows a learning process to arise.
A positive $\gamma$ can be seen as calibrating risk-seeking trait of the intelligent agent,
that is, the behavioral predispositions related to trading longer delays
for higher reward outcomes.
Such an agent puts relatively more emphasis on rewards expected in
a more distant future.
Concretely,
rewards that are not expected to come within
% \footnote{\textbf{N.B.:} $\frac{1}{1-\gamma} = 1 + \gamma + \gamma^2 + \gamma^3 + \ldots$}
$\tau := 1/(1 - \gamma)$
time steps from the present point are ignored.
The complexity reduction by time discounting
alleviates the variance of expected rewards accumulated across
considered action cascades by limiting the depth of the search tree.
Given that there is more uncertainty in the far future,
it is important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

@@

\subsubsection{Does the hippocampus subserve MCMC sampling?}
In RL, MCMC simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic}.
% \citep{silver2010monte}.
MCMC simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

In the human brain,
the HC could contribute to synthesizing imagined sequences of world states,
actions and rewards \citep{aronov2017, chao2017interaction, boyer2008evolutionary}.
These simulations of experience batches
would be used to update the value function, without ever looking inside the black box describing the model's dynamics.
A brain-imaging experiment in humans for instance identified hippocampal signals that specifically preceded upcoming choice performance in prospective planning in new environments
\citep{kaplan2017neural}.
This would be a simple control algorithm by evaluating all legal actions and selecting the action with
highest expected cumulative rewards.
In MDPs, MCMC simulation provides an effective mechanism both for tree search and for belief-based state updates, breaking the curse of dimensionality and allowing much greater scalability than has previously been possible \citep{silver2016mastering}.
This is because expected consequences of action choices can be well evaluated although only a subset of the states are actually considered \citep{daw2014algorithmic}.
% Much recent work points to MCMC or stochastic sampling-based approximations as a unifying framework for understanding how Bayesian inference may work practically across all these levels, in minds, brains, and machines. \textbf{XXX elvis: refs needed in last sentence. Plus, should be really bring up Bayesian?}

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state. This extension may partially account for implicit memory at the behavioral level, but may not explain the underlying neurobiological implementation or accommodate explicit memory. Implicit memory-based processing arises in our MDP account of DMN function in several different forms: successive updates of a) the action policy and the value function, both being products of the past, as well as b) the deep non-linear relationships within the hierarchical connections of biological neural networks. The brain's adaptive synaptic connections can be thought as a deep neural-network architecture affording an implicit form of information compression of life experience. Such memory traces are stored in the neural machinery and can be implicitly retrieved as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for probabilistic simulations across time-scales \citep{schacter2007remembering,Axelrod2017}.

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
activity when humans engage in cognitive processes that are relatively detached from
the current sensory environment.
The more familiar and predictable the current environment, the more brain resources may remain for allocating DMN activity to MDP processes extending beyond the present time and sensory context. In line with this perspective, DMN engagement was shown to heighten and relate to effective behavioral responses in the practiced phase of a demanding cognitive flexibility task, as compared to acquisition phase when participants learned context-specific rules. This involvement in automated decision-making has led the authors to propose an ``autopilot'' role for the DMN \citep{vatansever2017pnas}, which may contribute to optimizing control of the organism in general. Among all parts of the DMN, the RTPJ is perhaps the most evident candidate for a network-switching relay that calibrates between processing of environment-engaged versus internally generated information \citep{downar2000multimodal, golland2006extrinsic, bzdok2013tpj}.

@@

From the perspective of a model-free RL agent,
prediction in the DMN reduces to
generalization of
policy and value computations from sampled experiences to
successful action choices and reward predictions in future states.
As such,
plasticity in the DMN arises naturally.
If an agent behaving optimally in a certain environment moves
to new, yet unexperienced environment, reward prediction errors will
largely increase.
This feedback will lead to adaptation of policy considerations
and value estimations until the intelligent system converges to a
new steady state of optimal action decisions in a volatile world.

@@

\item Experiment (Humans): We hypothesize that the relevant time horizon is modulated by various factors such as age, acute stress and time-enduring impulsivity traits. Using a temporal discounting experiment, it can be quantified how the time horizon is affected at the behavioral level and then traced-back to its corresponding neural representation. Such experimental investigation can be designed to examine between-group and within-group effects (e.g., impulsive population like chronic gamblers or drug addicts); and brought in context with the participant’s age and personality traits.

@@

In sum,
recasting DMN function in terms of MDPs therefore naturally incorporates
a majority of aspects from the prediction coding hypothesis.
The present MDP account of DMN function may therefore
serve as a concrete implementation of
predictive coding ideas.
MDPs have the advantage of exposing an explicit
mechanisms for controlling the horizon of future considerations and for
how the internal representation of the world is updated,
as well as why certain predictions may be more relevant to the agent than others.

@@

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be explored.
In both accounts,
there is also little reason to assume that contemplating alternative realities of various
levels of complexity, abstraction, time-scale, and purpose
rely on mechanisms that are qualitatively different. This interpretation concurs with
DMN activity increases across time, space, and content domains
demonstrated in many brain-imaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account offer explanations why HC damage does
not only impair recalling past events, but also imagining hypothetical and future
scenarios \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-enabled, internally
generated information for probabilistic representation of action outcomes,
MDPs render explicit the grounds on which an action is
eventually chosen, namely, the estimated cumulative reward.
In contrast to many versions of the semantic hypothesis,
the MDPs naturally integrate the egocentric view
(more related to current action, state, and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not provide
suffcient explanation of \textit{how}
explicit world knowledge and logical analogies thereof
lead to foresight of future actions and states.
The semantic hypothesis does also not fully explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space exploration
by MCMC approximation achieves better fine-tuning of the
action policies and inference of expected reward outcomes.
That is, the purposeful stochasticity of policy and value updates
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, mental scene construction according to the semantic
account is lacking an explicit time and incentive model,
both of which are integral parts of the MDP interpretation of DMN function.
